---
title: "Group_06_Analysis"
author: "Group 06"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: 
   geometry: "left=2cm, right=2cm, top=2cm, bottom=2cm"
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

```{r}
#| label: libraries
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)
library(MASS)
library(knitr)
library(GGally)
```


# Introduction {#sec-Intro}


```{r}
# Import the data sets
imdb_data <- read.csv("dataset06.csv")
glimpse(imdb_data)
```

# Data Wrangling Methods {#sec-DW}
Before we begin the analysis of our data, let's transform the data using various tools.The process below describes the detailed data wrangling techniques that are used to get the desired data set.
```{r}
# Since genre is of type character, we convert it to factor
imdb_data$genre <- as.factor(imdb_data$genre)
```

As per the research question, 'rating' is converted to a binary variable taking 1 or 0 values.
```{r}
#new binary column#
imdb_data$high_rating <- ifelse(imdb_data$rating >= 7, 1, 0)
```

A check for missing values was conducted and it was found that 103 observation were missing from the column length. Impute missing values with the median since median is a robust measure i.e not impacted by outliers as much as mean
```{r}
# Check for missing values
missing_values <- sum(is.na(imdb_data$length)) #103 values
colSums(is.na(imdb_data)) #103 observations are missing in 'length'

# Impute missing values with the median 
median_length <- median(imdb_data$length, na.rm = TRUE)
imdb_data$length[is.na(imdb_data$length)] <- median_length
```

# Exploratory Data Analysis {#sec-EDA}
## scatterplot

Let's have a look at the first five rows of the data frame
```{r}
<<<<<<< Updated upstream
#| label: fig-plot1
#| fig-cap: scatterplot matrix with ggpairs
#| fig-align: center
#| message: false
#Add a new column#
imdb_data <- imdb_data %>%
  mutate(log_votes = log(votes))

imdb_data <- imdb_data %>%
  mutate(log_length = log(length))

ggpairs(imdb_data_numeric[, c("year", "budget", "rating","length","log_length","votes","log_votes")], 
                              title="Scatterplot matrix")
```

## property1:length(ln)

```{r}
#| label: fig-plot2
#| fig-cap: boxplot of length(ln)
#| fig-width: 4
#| fig-height: 5
#| fig-align: center
#| message: false
#Add a binary variable as a column
imdb_data$rate<- ifelse(imdb_data$rating >= 7, "high rating", "low rating")
#boxplot
ggplot(data = imdb_data, aes(x = rate, y = log_length,fill=rate) )+
  geom_boxplot() +
  labs(x = "rating", y = "log_length")
```

@fig-plot2 shows that the median film log_length of high rating films is less than that of low rating films. Interquartile Range of low rating is smaller than that of high rating, which shows that the data about low rating is more concentrated.

```{r}
#| label: table1
#| tbl-cap: Summary statistics on length.
#Add a column to make the table clearer
table=imdb_data %>%
  group_by(rate) %>%
  summarise('Mean' = mean(log_length),
            'Median' = median(log_length),
            'St.Dev' = sd(log_length),
            'Min' = min(log_length),
            'Max' = max(log_length),
            'IQR' = quantile(log_length,0.75)-quantile(log_length,0.25),
            'Sample_size' = n())
table|>
  gt() |>
  fmt_number(decimals=2) |>
  cols_label(
    rate=html("rate"),
    Mean = html("Mean"),
    Median = html("Median"),
    St.Dev = html("Std. Dev"),
    Min = html("Minimum"),
    Max = html("Maximum"),
    IQR = html("Interquartile Range"),
    Sample_size = html("Sample Size")
  )
```

@table1 shows that the size of the film with low rating is more than twice as many as that with high rating. The mean log_length film with high rating(3.62) is lower than that with low rating(4.52).On the whole, highly rated films have a low parity of length phrases. But a low rated films is more stable.

## property2:budget(in \$1000000s)

```{r}
#| label: fig-plot3
#| fig-cap: boxplot of budget
#| fig-width: 4
#| fig-height: 5
#| fig-align: center
#| message: false
#boxplot
ggplot(data = imdb_data, aes(x = rate, y = budget,fill=rate) )+
  geom_boxplot() +
  labs(x = "rating", y = "budget")
```

@fig-plot3 shows that the median budget film of high rating films is higher than that of low rating films.

```{r}
#| label: table2
#| tbl-cap: Summary statistics on budget.
table=imdb_data %>%
  group_by(rate) %>%
  summarise('Mean' = mean(budget),
            'Median' = median(budget),
            'St.Dev' = sd(budget),
            'Min' = min(budget),
            'Max' = max(budget),
            'IQR' = quantile(budget,0.75)-quantile(budget,0.25),
            'Sample_size' = n())

table|>
  gt() |>
  fmt_number(decimals=2) |>
  cols_label(
    Median = html("Median"),
    St.Dev = html("Std. Dev"),
    Min = html("Minimum"),
    Max = html("Maximum"),
    IQR = html("Interquartile Range"),
    Sample_size = html("Sample Size")
  )
```

@table2 shows that the mean budget film with high rating(13.09 in \$1000000s) is higher than that with low rating(11.51 in \$1000000s).And the median budget film with high rating(13.00 in \$1000000s) is higher than that with low rating(11.50 in \$1000000s).In general, films with high rating have higher budgets than movies with low rating. ##votes

## property3: votes(ln)

```{r}
#| label: fig-plot4
#| fig-cap: boxplot of votes(ln)
#| fig-width: 4
#| fig-height: 5
#| fig-align: center
#| message: false
#boxplot
ggplot(data = imdb_data, aes(x = rate, y = log_votes,fill=rate) )+
  geom_boxplot() +
  labs(x = "rating", y = "log_votes")
```

@fig-plot4 shows that the median log_votes film of high rating films is lower than that of low rating films.

```{r}
#| label: table3
#| tbl-cap: Summary statistics on votes(ln).
table=imdb_data %>%
  group_by(rate) %>%
  summarise('Mean' = mean(log_votes),
            'Median' = median(log_votes),
            'St.Dev' = sd(log_votes),
            'Min' = min(log_votes),
            'Max' = max(log_votes),
            'IQR' = quantile(log_votes,0.75)-quantile(log_votes,0.25),
            'Sample_size' = n())

table|>
  gt() |>
  fmt_number(decimals=2) |>
  cols_label(
    Mean = html("Mean"),
    Median = html("Median"),
    St.Dev = html("Std. Dev"),
    Min = html("Minimum"),
    Max = html("Maximum"),
    IQR = html("Interquartile Range"),
    Sample_size = html("Sample Size")
  )
```

@table3 shows that the mean log_votes film with high rating(3.47) is lower than that with low rating(4.03). In general, there were fewer log_votes for films with high ratings.

## property4:genre

```{r}
library(janitor)
imdb_data %>% 
  tabyl(genre, rate) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>%
adorn_ns() # To show original counts
```

```{r}
#| label: fig-plot5
#| fig-cap: bar charts of genre
#| fig-width: 4
#| fig-height: 5
#| fig-align: center
#| message: false
g <- ggplot(imdb_data, aes(genre))
# Number of cars in each class:
g + geom_bar(aes(fill = rate))
```

@fig-plot5 shows that the number of low rating films is bigger than the number of high rating films in the genre films about Action,Drama,Romance.Besides,the number of low rating films is smaller than the number of high rating films in the genre films about Animation,Documentary,Short. In genre Comedy, the number of films with high ratings is about the same as the number of films with low scores.
=======
imdb_data |>
  slice_head(n=5) |>
  gt() |>
  fmt_number(decimals=2)
```

## Summary Statistics {#sec-sum}
```{r}
summary(imdb_data)
```

## Correlation {#sec-cor}
Scatter plot Matrix shows very low correlation between categorical variables
```{r}
scatterplot_matrix <- ggpairs(imdb_data[, c("year", "budget", "rating", "length", "votes")], 
                              title="Scatterplot matrix")
scatterplot_matrix
```
## Visualization {#sec-viz}

Histograms to understand the data structures of different variables

```{r}
#DATA STRUCTURES#
#Histograms for statistical distribution#
year_plot <- ggplot(data=imdb_data, mapping=aes(x=year))+
  geom_histogram(color="red")

length_plot <- ggplot(data=imdb_data, mapping=aes(x=length))+
  geom_histogram(color="blue")

budget_plot <- ggplot(data=imdb_data, mapping=aes(x=budget))+
  geom_histogram(color="green")

votes_plot <- ggplot(data=imdb_data, mapping=aes(x=votes))+
  geom_histogram(color="yellow")

log_votes_plot <- ggplot(data=imdb_data, mapping=aes(x=log(votes)))+
  geom_histogram(color="yellow")

rating_plot <- ggplot(data=imdb_data, mapping=aes(x=rating))+
  geom_histogram(color="skyblue")

grid.arrange(year_plot,budget_plot,votes_plot, log_votes_plot,
             length_plot,rating_plot, ncol=2)
```

Based on the results above the following transformations are done
```{r}
#Add a new column#
imdb_data <- imdb_data %>%
  mutate(log_votes = log(votes))
```

Scatterplots to understand the relationship of Response variable with numerical explanatory variables
```{r}
s1 <- ggplot(data=imdb_data, aes(x = year, y = rating, color=genre))+
  geom_point()

s2 <- ggplot(data=imdb_data, aes(x = length, y = rating, color=genre))+
  geom_point()

s3 <- ggplot(data=imdb_data, aes(x = budget, y = rating, color=genre))+
  geom_point()

s4 <- ggplot(data=imdb_data, aes(x = log_votes , y = rating, color=genre))+
  geom_point()
grid.arrange(s1, s2, s3, s4, ncol=2)
```


## Boxplots {#sec-box}

The boxplots are created to understand various numerical variables individual and to explore if the categorical variable 'genre' has an impact on it
```{r}
year_outliers <- imdb_data %>%
  group_by(genre) %>%
  mutate(is_outlier = year > quantile(year, 0.75) + 1.5 * IQR(year) |
           year < quantile(year, 0.25) - 1.5 * IQR(year)) %>%
  filter(is_outlier) %>%
  ungroup() # 14 outliers
p1 <- ggplot(imdb_data, aes(x = genre, y = year))+
  geom_boxplot(fill=c("red", "green", "yellow", "blue", "pink", "white", "lightblue"))
p1
```
Comment on boxplot above median along with outliers

```{r}
length_outliers <- imdb_data %>%
  group_by(genre) %>%
  mutate(is_outlier = length > quantile(length, 0.75) + 1.5 * IQR(length) |
           length < quantile(length, 0.25) - 1.5 * IQR(length)) %>%
  filter(is_outlier) %>%
  ungroup() #207 outliers with length
p2 <- ggplot(imdb_data, aes(x = genre, y = length))+
  geom_boxplot(fill=c("red", "green", "yellow", "blue", "pink", "white", "lightblue"))
p2
```
Comment on boxplot above median along with outliers


```{r}
p3 <- ggplot(imdb_data, aes(x = genre, y = budget))+
  geom_boxplot(fill=c("red", "green", "yellow", "blue", "pink", "white", "lightblue"))
p3
budget_outliers <- imdb_data %>%
  mutate(is_outlier = budget > quantile(budget, 0.75) + 1.5 * IQR(budget) |
           budget < quantile(budget, 0.25) - 1.5 * IQR(budget)) %>%
  filter(is_outlier) %>%
  ungroup() # 16 outliers
```
Comment on boxplot above median along with outliers

```{r}
log_votes_outliers <- imdb_data %>%
  group_by(genre) %>%
  mutate(is_outlier = log_votes > quantile(log_votes, 0.75) + 1.5 * IQR(log_votes) |
           log_votes < quantile(log_votes, 0.25) - 1.5 * IQR(log_votes)) %>%
  filter(is_outlier) %>%
  ungroup() # 294 outliers with votes, 57 outliers with log_votes

p4 <- ggplot(imdb_data, aes(x = genre, y = log_votes))+
  geom_boxplot(fill=c("red", "green", "yellow", "blue", "pink", "white", "lightblue"))
p4
```

```{r}
rating_outliers <- imdb_data %>%
  group_by(genre) %>%
  mutate(is_outlier = rating > quantile(rating, 0.75) + 1.5 * IQR(rating) |
           rating < quantile(rating, 0.25) - 1.5 * IQR(rating)) %>%
  filter(is_outlier) %>%
  ungroup() #138 outliers
p5 <- ggplot(imdb_data, aes(x = genre, y = rating))+
  geom_boxplot(fill=c("red", "green", "yellow", "blue", "pink", "white", "lightblue"))
p5
```


>>>>>>> Stashed changes

# Formal Data Analysis {#sec-FDA}

# Conclusions {#sec-Conc}
