---
title: "Group_06_Analysis"
author: "Group 06"
number-sections: true
format: 
  html:
    embed-resources: true
    code-tools: true
  pdf: default
editor_options: 
  chunk_output_type: console
execute:
  echo: false
  eval: true
  warning: false
  message: false
---

```{r}
#| label: libraries
library(ggplot2)
library(tidyverse)
library(gt)
library(patchwork)
library(gridExtra)
library(moderndive)
library(MASS)
library(knitr)
library(GGally)
```


# Introduction {#sec-Intro}


```{r}
# Import the data sets
imdb_data <- read.csv("dataset06.csv")
glimpse(imdb_data)
```

# Data Wrangling Methods {#sec-DW}
Before we begin the analysis of our data, let's transform the data using various tools.The process below describes the detailed data wrangling techniques that are used to get the desired data set.
```{r}
# Since genre is of type character, we convert it to factor
imdb_data$genre <- as.factor(imdb_data$genre)
```

As per the research question, 'rating' is converted to a binary variable taking 1 or 0 values.
```{r}
#new binary column#
imdb_data$high_rating <- ifelse(imdb_data$rating >= 7, 1, 0)
```

A check for missing values was conducted and it was found that 103 observation were missing from the column length. Impute missing values with the median since median is a robust measure i.e not impacted by outliers as much as mean
```{r}
# Check for missing values
missing_values <- sum(is.na(imdb_data$length)) #103 values
colSums(is.na(imdb_data)) #103 observations are missing in 'length'

# Impute missing values with the median 
median_length <- median(imdb_data$length, na.rm = TRUE)
imdb_data$length[is.na(imdb_data$length)] <- median_length
```

# Exploratory Data Analysis {#sec-EDA}

```{r}

```


# Formal Data Analysis {#sec-FDA}

# Conclusions {#sec-Conc}
